---
format:
  html:
    toc: false
    embed-resources: true
---

{{< include problem_statement.qmd >}}

```{r}
#| echo: false
#| output: false
library(tidyverse)
library(kableExtra)
library(knitr)
results <- read.csv("reb_4_4_data.csv")
```

```{r}
#| echo: false
kbl(results, format = "markdown", align = "cc", col.names = c("i (Expt. Number)", "T (°C)",
    "k (L mol^-1^ min^-1^)"))
```

**Necessary Equations**

$$
k = k_0 \exp{\left(  \frac{-E}{RT} \right)} \tag{1}
$$

$$
\beta = \log_{10} {\left(k_0\right)} \tag{2}
$$

$$
k = 10^\beta \exp{\left(  \frac{-E}{RT} \right)} \tag{3}
$$

$$
T_{K,i} = T_i + 273.15 \tag{4}
$$

$$
\beta = \log_{10} {\left(k_0\right)} \quad \Rightarrow \quad k_0 = 10^\beta \tag{5}
$$

$$
k_{i,p} = k_0 \exp{\left(  \frac{-E}{RT_{K,i}} \right)} \tag{6}
$$

$$
\epsilon_i = k_i - k_{i,p} \tag{7}
$$

**Calculations**

To use non-linear parameter estimation software:

1. Calculate the absolute temperature for each data point use them to construct the adjusted input vector, $\underline{x}$ using equation (4).
2. Generate the experimental response vector, $\underline{y_{expt}}$, using equation (5).
3. Fit equation (3) to the resulting $\underline{x}$ - $\underline{y_{expt}}$ data using parameter estimation software.
4. Calculate $k_{0,j}$ from the best estimate for $\beta$, using equation (6).
5. Generate the predicted response vector, $\underline{y_{pred}}$, using equation (7).
6. Generate the residuals vector, $\underline{\epsilon}$, using equation (8).
7. Generate a parity plot by plotting $y_{expt,i}$ *vs*. $y_{pred,i}$}.
8. Generate a residuals plot by plotting $\epsilon_i$ *vs*. $x_i$.
6. Assess the accuracy of the resulting model using the following criteria for an accurate model:
    a. The coefficient of determination, $R^2$, is close to 1.
    b. The uncertainty in each model parameter is small compared to its value.
    c. The points in the parity plot are all be close to a diagonal line passing through the origin.
    d. In the residuals plot, the points scatter randomly about the horizontal axis, and no systematic deviations are apparent.

To use linear least squares software or a spreadsheet program:
