---
format:
  html:
    toc: false
    embed-resources: true
---

{{< include problem_statement.qmd >}}

```{r}
#| echo: false
#| output: false
library(tidyverse)
library(kableExtra)
library(knitr)
results <- read.csv("reb_4_4_data.csv")
```

```{r}
#| echo: false
kbl(results, format = "markdown", align = "cc", col.names = c("Temperature (°C)",
    "Rate Coefficient (L mol^-1^ min^-1^)"))
```

**Necessary Equations**

$$
k_j = k_{0,j} \exp{\left(  \frac{-E_j}{RT} \right)} \tag{1}
$$

$$
\ln{k_j} = \left(  \frac{-E_j}{R} \right)\left(  \frac{1}{T} \right) + \ln{k_{0,j}} \tag{2}
$$

$$
y = mx + b  \tag{3}
$$

$$
y = \ln{k_j} \tag{4}
$$

$$
x = \frac{1}{T} \tag{5}
$$

$$
m = \frac{-E_j}{R} \tag{6}
$$

$$
b = \ln{k_{0,j}} \tag{7}
$$

**Calculations**

To use linear least squares software or a spreadsheet program:

1. Calculate the absolute temperature for each data point.
2. Calculate $y$ and $x$ for each data point using equations (4) and (5).
3. Fit equation (3) to the resulting data
4. Calculate $E_j$ from the resulting slope, $m$, using equation (6).
5. Calculate $k_{0,j}$ from the resulting intercept, $b$, using equation (7).
6. Assess the accuracy of the resulting model.
    a. Plot equation (3) and the experimental $x$, $y$ data on a common set of axes; the data points will all be close to the line if the fit is accurate, and there won't be any systematic differences between the data and the line.
    b. The more accurate the model, the closer the coefficient of determination, $R^2$, will be to 1.0.
    c. The estimated uncertainties in the parameter values will be small relative to the corresponding parameter value when the model is accurate.

To use non-linear parameter estimation software:

1. Create a vector containing the experimental values of the temperature in K.
2. Create a vector containing the measured values values of the rate coefficient.
3. Create a function or subroutine that receives test values for the kinetics parameters, $k_{0,j}$ and $E_j$, along with the array of temperatures from step 1, calculates the value of the rate coefficient for each temperature using equation (1), and returns a vector containing the values of the rate coefficient at those temperatures.
4. Provide the results from steps 1 through 3 to a parameter estimation function along with a guess for the values of the kinetics parameters.